---
title: "Decathlon"
output: html_document
date: "2023-04-08"
---

# load libraries and read in data
```{r}
library(tidyverse)
library(here)
```

```{r}
decathlon_data <- read_csv(here("data/decathlon_clean.csv"))
```
```{r}
# view data and explore
glimpse(decathlon_data)
```

Q1: Who had the longest long jump seen in the data?

```{r}
decathlon_data %>%
  arrange(desc(long_jump)) %>%
  slice_max(long_jump, n = 1)
```

Q2: What was the average 100m time in each competition?

```{r}
decathlon_data %>% 
  group_by(competition) %>% 
  summarise(ave_100m = mean(run_100m))
```


Q3: Who had the highest total points across both competitions?
```{r}
decathlon_data %>%
  select(name, points, competition) %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points)) %>% 
  arrange(desc(total_points)) %>% 
  slice_max(total_points)

```

Q4: What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_data %>% 
  select(name, shot_put, competition) %>% 
  group_by(competition) %>% 
  arrange(desc(shot_put)) %>% 
  slice_max(shot_put, n = 3)

```

Q5: What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_data %>% 
  select(run_400m) %>% 
  mutate(less_than_50sec = run_400m < 50) %>% 
  group_by(less_than_50sec) %>% 
  summarise(ave = mean(run_400m))
```



