---
title: "Decathlon"
output: html_document
date: "2023-04-08"
---
# Introduction to project
The dataset contains athletics data from individuals competiting in two competitions (Olympics and Decastar) and sports such as long jump, 100m, hurdles and highjump. 
Some of the individuals have competed in both competitions. 

# Cleaning steps taken -
- cleaned up column headings to standard format
- the data has competitors names as rownames, edit these so that there is a new column for "name" and a separate rowname list
- renamed some of the column headings so it is easier to read
- write the cleaned data to a csv file to import for analysis



# load libraries and read in data
```{r}
library(tidyverse)
library(here)
```

```{r}
decathlon_data <- read_csv(here("data/decathlon_clean.csv"))
```




```{r}
# view data and explore
glimpse(decathlon_data)
```

Q1: Who had the longest long jump seen in the data?
Answer: Clay with 7.96m
```{r}
decathlon_data %>%
  arrange(desc(long_jump)) %>%
  slice_max(long_jump, n = 1)
```

Q2: What was the average 100m time in each competition?
Decastar = 11.17 seconds, Olympics = 10.9 seconds
```{r}
decathlon_data %>% 
  group_by(competition) %>% 
  summarise(ave_100m = mean(run_100m))
```


Q3: Who had the highest total points across both competitions?
Sebrle had total points of 17110
```{r}
decathlon_data %>%
  select(name, points, competition) %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points)) %>% 
  arrange(desc(total_points)) %>% 
  slice_max(total_points)

```

Q4: What was the shot-put scores for the top three competitors in each competition?
Answer: 
Decastar top three were Yurkov with 15.19, Sebrle with 14.83 and Karpov with 14.77
OlympicG top three were Sebrle with 16.36, Karpov with 15.93 and Macey with 15.73


```{r}
decathlon_data %>% 
  select(name, shot_put, competition) %>% 
  group_by(competition) %>% 
  arrange(desc(shot_put)) %>% 
  slice_max(shot_put, n = 3)

```

Q5: What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
Average points for competitors faster than 50sec = 7727, and with those slower than 50 secs = 8120

```{r}
decathlon_data %>% 
  select(run_400m, points) %>% 
  mutate(less_than_50sec = run_400m < 50) %>% 
  group_by(less_than_50sec) %>% 
  summarise(ave = mean(points))
```

